ohun: an R package for diagnosing and optimizing automated sound event detection

### 2022-08-16 manuscript/manuscript.Rmd
update manuscript

---
title: 'ohun: an R package for diagnosing and optimizing automatic sound event detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
Grace Smith-Vidaurre ^4,^ ^5,^ ^6^, 
Gloriana Chaverri ^3^, ^7^
Juan Carlos Brenes ^1^ &
Fabiola Chirino ^1^ &
Alejandro Rico-Guevara ^8,^ ^9^

@@

The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions. However, this diversity of automated detection tools also posits a challenge as it can be difficult to navigate [@Stowell2022]. In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform researchers’ decisions about which method better fits a given question and study system [@knight2017]. The performance of automated sound event detection routines have typically been evaluated using standard indices from signal detection theory [@balantic2020; @knight2017]. In its basic form, performance is assessed by comparing the output of a detection routine against a 'gold standard' reference in which all the target sounds have been annotated (hereafter called 'reference annotation'). This comparison facilitates quantifying the number of correctly detected sounds (true positives), wrongly detected sounds (false positives) and missed sounds (false negatives) as well as additional metrics derived from these indices (e.g. recall, precision). However, the fact that sound events are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures), but are rather embedded within a continuous string of sound, demand additional information to fully diagnose the temporal precision detection performance. This is particularly relevant when the precise time position of sounds is needed as is common when measuring the acoustic structure of sounds is the main goal [@Araya-Salas2017f]. For instance, the same signal can be detected as several separated sounds, the inferred time position can be offset from the the target signal position, or several sounds can be detected as a single one. Therefore, metrics that account for these additional performance dimensions are a valuable tool to properly diagnose automatic sound event detection.

@@

The *ohun* package makes use of signal detection theory indices to evaluate detection performance. Signal detection theory deals with the process of recovering sounds (*i.e.* target sounds) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty [@hossin2015]. During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target sounds correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, sounds incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise). However, TNs cannot always be easily defined in the context of sound event detection as noise cannot always be partitioned in discrete units. Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine and are widely used in the context of sound event detection [@knight2017]: recall (correct detections relative to total detections), precision (proportion of target sounds that were correctly detected) and F1 score (combined recall and precision as the harmonic mean of these two, providing a single value for evaluating performance, a.k.a. F-score, F-measure or Dice similarity coefficient).

>>>>

---
title: 'ohun: an R package for diagnosing and optimizing automated sound event detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
Grace Smith-Vidaurre ^4,^ ^5,^ ^6^, 
Gloriana Chaverri ^3^, ^7^
Juan Carlos Brenes ^1^
Fabiola Chirino ^1^ &
Alejandro Rico-Guevara ^8,^ ^9^

@@

The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions. However, this diversity of automated detection tools also posits a challenge as it can be difficult to navigate [@Stowell2022]. In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform researchers’ decisions about which method better fits a given question and study system [@knight2017]. The performance of automated sound event detection routines have typically been evaluated using standard indices from signal detection theory [@macmillan2004; @balantic2020; @knight2017]. In its basic form, performance is assessed by comparing the output of a detection routine against a 'gold standard' reference in which all the target sounds have been annotated (hereafter called 'reference annotation'). This comparison facilitates quantifying the number of correctly detected sounds (true positives), wrongly detected sounds (false positives) and missed sounds (false negatives) as well as additional metrics derived from these indices (e.g. recall, precision). However, the fact that sound events are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures), but are rather embedded within a continuous string of sound, demand additional information to fully diagnose the temporal precision detection performance. This is particularly relevant when the precise time position of sounds is needed as is common when measuring the acoustic structure of sounds is the main goal [@Araya-Salas2017f]. For instance, the same signal can be detected as several separated sounds, the inferred time position can be offset from the the target signal position, or several sounds can be detected as a single one. Therefore, metrics that account for these additional performance dimensions are a valuable tool to properly diagnose automatic sound event detection.

@@

The *ohun* package makes use of signal detection theory indices to evaluate detection performance. Signal detection theory deals with the process of recovering sounds (*i.e.* target sounds) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty [@hossin2015]. During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target sounds correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, sounds incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise). However, TNs cannot always be easily defined in the context of sound event detection as noise might not be partitioned in discrete units. Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine and are widely used in the context of sound event detection [@knight2017]: recall (correct detections relative to total detections), precision (proportion of target sounds that were correctly detected) and F1 score (combined recall and precision as the harmonic mean of these two, providing a single value for evaluating performance, a.k.a. F-score, F-measure or Dice similarity coefficient).
